\ttitle

### 2014-01-17 Appendices/AppendixA-Results.tex
Move results Chapter 6 to Appendix'

% !TEX root = ../../main.tex
\section{Results}\label{sec:real_world_results}
In this section we will provide the results of our methods \gls{ocs-hats} applied to the real-world data sets and discuss the quality of the solution.
As stated above, we will use the objective quality metrics as used in \Cref{sec:artificial_data_quality_metrics}.
That will be in the form of a tabular overview of the (manually set) best meta-parameters and results, expressed in \gls{far} $\gamma$ and average delay $d_{avg}$ (following \Cref{sec:artificial_data_results}).
Furthermore, we will discuss the results from an empirical point of view, since the problem of finding change points in human activities is ambiguous.
By stating our observations and remarks and providing fragments of visual results, we will discuss the characterizing strengths and weaknesses of our method.

\subsection{Objective metrics}
The employed two-stage algorithm requires the setting of user defined parameters.
In \Cref{tab:results_real_world} for each run the results are listed, together with the used parameters.
We have used a window length of 50 frames, which represents around $1.16$ seconds of data (the frame-rate of the recordings is not fixed).
The other parameter values, \ie the \gls{rbf} $\sigma$, high and low threshold, and proximity parameter $\delta$, are the result of an educated guess and manual optimization.
In general, a high value for the upper threshold and low value for the lower threshold indicates a robust method, since it indicates that change points are significantly different from homogeneous segments.
The proximity $\delta$ should be low, since that indicates there are few false positives.

@@

To give a better impression on the quality of the $\operatorname*{Average\_delay}$, \Cref{fig:boxplot_real_world_runs} shows a box plot for each run.
It shows the spread of the $\operatorname*{Average\_delay}$ in seconds over all the change points.
A lower and more compact box plot is better.

\begin{figure}
\centering
  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/boxplot_results_real_world_runs.eps}
  \caption[Box plot results real-world runs]{Box plot of the results for the real-world runs, indicating the number of data points between the actual and closest detected change points. A lower and more compact box plot is better.}
  \label{fig:boxplot_real_world_runs}
\end{figure}

\subsection{Subjective observations and remarks}
It can be ambiguous where to place the (manually annotated) change points, in the case of real-world data.
The objective quality measure can give a distorted view on the performance.
To overcome this, we will state observations and remarks found after close (visual) inspection of the annotated and discovered change points.
In \Cref{fig:plots_all_runs_results} all the results are plotted.
The black solid lines indicate our manually determined change points.
The dashed purple lines indicate the change points as discovered by our method.
We will start with general observations applicable to all the runs and will continue with individual remarks for each run.
Where possible, a graphical illustration will support our observations.

Overall our impression is that the proposed method is fairly good at recognizing change points.
Depending on the parameters for the sensitivity adjustment, almost all changes are discovered.
With a high sensitivity it seems, initially, the method produces many false positives.
After closer inspecting of the data, we can confirm that there is a change in the data, although the performed activity stayed the same.
This is also a weakness of the method: since the sensitivity is set globally, it can produce many semi-false positive change points for some segments of a run.

An other positive observation is that the method often recognizes a change point \emph{before} our annotation.
Again, after closer inspection we can see that the data changes shape and distribution before we annotated a change based on the video recordings.
This effect can be observed in the case of a transition from running to walking; the last `steps' of running are already recognized of being different.
As with the above behavior, this is subject to parameter settings.

Considering the overall performance and results, we have the following observations of cases in which the method has problems with detecting change:
\begin{itemize}
  \item \textbf{Merging:} in our method we use a proximity time period $\delta$ (usually up to $0.5$ - $1.0$ seconds) to merge discovered change points that have a very small distance between them.
  Multiple merging-strategies can be applied, and in our method we use the naive implementation by simply ignoring all the discovered change points that occur less than $\delta$ seconds after the previous change point.
  This works well when there is a block of noisy data with a high amount of (falsely) discovered change points.
  On the flip side, this method will ignore real new change points which occur during the noisy period.
  This problem on itself can be formulated as a change detection problem.
  \item \textbf{Masking:} Since the parameters for the detection method are set globally, it is difficult to discover all the change points (and only the change points) without a high \gls{far}.
  If a change point is proceeded by a noisy block, then the real change point will be merged with the previous change points.
  In an other case, where a change point is very clear represented in the data but the next, close by, other change point is more subtle, the latter change point is \emph{masked} by the former.
  This masking effect is also discussed in \cite{inclan1994use}, where an iterated approach is applied.
  \item \textbf{False heterogeneity:} For our video-data synchronization, we started and ended each recording with a few seconds in which the recording smartphone was kept still in the air.
  During this period, the data variance, and thus the constructed hypersphere, becomes very small.
  As a results, even small movements are considered to be changes and in the case of the mid-air still smartphone a lot of false change points are detected.
  Due to the merging effect described above, the final real change point (often shaking the smartphone) is not discovered.
  \item \textbf{Incorrect weighting:} In our analyses we have used the data from the accelerometer, magnetic field, and rotation sensors.
  In the segments which embodied movement in a circular manner (such as walking and running around a fountain), the turn was not (or at a different time point) discovered.
  Alternatively, when we only used the magnetic field and rotation sensors, the turn was correctly discovered.
  Other transitions, such as from walking to running, are harder to discover without the (linear) accelerometer sensor data.
\end{itemize}

When looking at the individual runs, we have the following observations for each run number:
\begin{enumerate}
  \item \textbf{Subject 2, straight walk and run} \Cref{fig:data_with_cps_run_1}
    \begin{itemize}
      \item The first transition from running to walking, around $14s$ is harder to discover than the transition for the same activities around $33s$.
      This shows is that in real-world applications there is a diversity between the same transitions and activities.
      \item Around $8s$, $9s$, and $16s$ a few steps (while walking) are regarded as change points.
      During the running segments from $17s$ and $28s$ there is a lower probability of change for each step.
    \end{itemize}
  \item \textbf{Subject 1, straight walk and run} \Cref{fig:data_with_cps_run_2}
    \begin{itemize}
      \item Following the video recordings, we have annotated a change point from running to walking around $37s$.
      Our method discovers a change point almost a second before.
      In retrospect, we can see that the data distribution indeed changes from the discovered change point on.
      This shows us two important principles.
      The first is that the annotations are very subjective.
      The second is that between different activities the transition period is longer than we would think.
      Looking at the data, we can see that the body slows down, even before we visually notice it on the video recordings.
    \end{itemize}

@@

\section{Possible improvements}\label{sec:possible_improvements}
From the above section we can extract a few possible improvements for the algorithm.
Most of the observations and remarks are due to a different required sensitivity for local segments of the run.
Currently, our method employs a global setting for the thresholds.
A possible improvement would be to use locally optimized parameters, although that would require an reflective data processing approach or more a priori information about the data.
With our approach, the lack of a priori knowledge about the data distribution is one of the design decisions.

Furthermore, we discovered that sometimes, especially with movements incorporating circular motions, the addition of (linear) accelerometer data makes the discovery of change points harder.
In such cases, each of the inertial sensor data streams should be relatively weighted (\eg using the covariance) in order to signal an overall change when it only occurs in one of the streams.

>>>>

% !TEX root = ../main.tex
% Appendix A

\chapter{Observations for Real-World Results} % Main appendix title

\label{AppendixA} % For referencing this appendix elsewhere, use~\ref{AppendixA}

\lhead{Appendix A. \emph{Real-World Observations}} % This is for the header on each page - perhaps a shortened title

@@

\begin{figure}
\centering
  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/boxplot_results_real_world_runs.eps}
  \caption[Box plot results real-world runs individually optimized]{Box plot of the results for the real-world runs, indicating the number of data points between the actual and closest detected change points. A lower and more compact box plot is better.}
  \label{fig:boxplot_real_world_runs}
\end{figure}

This appendix provides observations and remarks which we made from inspecting the real-world results from \Cref{Chapter6}.
In \Cref{subsec:subjective_results} we have discussed a few overall effects.
In the following we discuss the performance of each run, using the (manually determined) specific parameters as listed in \Cref{tab:results_real_world}.

To give a better impression on the quality of the $d_{avg}$, \Cref{fig:boxplot_real_world_runs} shows a box plot for each run.
It shows the spread of the $d_{avg}$ in seconds over all the change points.
A lower and more compact box plot is better.

When looking at the individual runs, we have the following observations for each run:

\begin{enumerate}
  \item \textbf{Subject 2, straight walk and run} \Cref{fig:data_with_cps_run_1}
    \begin{itemize}
      \item The first transition from running to walking, around $14s$ is harder to discover than the transition for the same activities around $33s$.
      This shows is that in real-world applications there is a diversity between the same transitions and activities.
      \item Around $8s$, $9s$, and $16s$ a few steps (while walking) are regarded as change points.
      During the running segments from $17s$ and $28s$ there is a lower probability of change for each step.
    \end{itemize}
  \item \textbf{Subject 1, straight walk and run} \Cref{fig:data_with_cps_run_2}
    \begin{itemize}
      \item Following the video recordings, we have annotated a change point from running to walking around $37s$.
      Our method discovers a change point almost a second before.
      In retrospect, we can see that the data distribution indeed changes from the discovered change point on.
      This shows us two important principles.
      The first is that the annotations are very subjective.
      The second is that between different activities the transition period is longer than we would think.
      Looking at the data, we can see that the body slows down, even before we visually notice it on the video recordings.
    \end{itemize}

@@

### 2014-01-17 Chapters/Chapter1.tex
Move results Chapter 6 to Appendix'

\section{Thesis structure}\label{sec:intro_thesis_structure}
The structure of this thesis is as follows.
In \Cref{Chapter2} a literature review is provided.
We will look at the different interpretations and implementations for change, novelty, and outlier detection.
Previous work in the field of \gls{har} is discussed and we end with existing applications of \glspl{svm} to change detection.
\Cref{Chapter3} further analyses \glspl{svm} and two implementations of \glspl{oc-svm}.
It relates the properties of the constructed \gls{oc-svm} models to change detection in time series data.
That relation is applied to our problem formulation in \Cref{Chapter4}, where we construct our change detection method \TODO{add name}, based on the work of Camci~\cite{camci2010change}.
In \Cref{Chapter5} we apply the method to artificial data and show the objective performance.
In \Cref{Chapter6} the method is applied to our real-world \gls{almende-data} \cite{vlasveld2014acras}.
In that chapter we show the ability of detecting change in \gls{har} data, recorded by inertial sensors.
This thesis is concluded in \Cref{Chapter7}, in which we reflect on the performed research and state possibilities for feature research.

>>>>

\section{Thesis structure}\label{sec:intro_thesis_structure}
The structure of this thesis is as follows.
In \Cref{Chapter2} a literature review is provided.
We will look at the different interpretations and implementations for change, novelty, and outlier detection.
Previous work in the field of \gls{har} is discussed and we end with existing applications of \glspl{svm} to change detection.
\Cref{Chapter3} further analyses \glspl{svm} and two implementations of \glspl{oc-svm}.
It relates the properties of the constructed \gls{oc-svm} models to change detection in time series data.
That relation is applied to our problem formulation in \Cref{Chapter4}, where we construct our change detection method \gls{ocs-hats}~\cite{vlasveld2014hats}, based on the work of Camci~\cite{camci2010change}.
In \Cref{Chapter5} we apply the method to artificial data and show the objective performance.
In \Cref{Chapter6} the method is applied to our real-world \gls{almende-data} \cite{vlasveld2014acras}.
In that chapter we show the ability of detecting change in \gls{har} data, recorded by inertial sensors.
This thesis is concluded in \Cref{Chapter7}, in which we reflect on the performed research and state possibilities for feature research.

### 2014-01-17 Chapters/Chapter4.tex
Move results Chapter 6 to Appendix'

This chapter will introduce our method, which we named \gls{ocs-hats}, and setup for the temporal segmentation of human activities, using an \gls{oc-svm} approach.
It shows our method to apply a \gls{oc-svm} based temporal segmentation method to inertial sensor time series data, often used in the context of \acrlong{har}.
As far as we know it is the first in its kind, especially on the application of \gls{svdd} and \gls{svcpd} to real-world \gls{har} data.
The setup follows the structure as illustrated in \Cref{fig:method_overview} and further discussed in \Cref{sec:method_overview}.
It starts with the data collection phase, in which we both use artificial and real-world data.
The data sets are discussed in \Cref{sec:data_gathering}.
After (optional) pre-processing, the \gls{svdd} algorithm is used to create a model representation, which is iteratively updated.
In \Cref{sec:method_model_construction} this construction phase is discussed and shows the windowed approach.
From this constructed model we can extract some features.
In our case, we use the approximated radius $R$ of the hypersphere to detect changes.
\Cref{sec:method_model_features} goes into detail about the rationale of using that feature.
Finally, in \Cref{sec:method_change_detection} we discuss two different, but simple, methods to detect changes in the time series data, based on the extracted radius $R$.

>>>>

This chapter will introduce our method, which we named \gls{ocs-hats}~\cite{vlasveld2014hats}, and setup for the temporal segmentation of human activities, using an \gls{oc-svm} approach.
It shows our method to apply a \gls{oc-svm} based temporal segmentation method to inertial sensor time series data, often used in the context of \acrlong{har}.
As far as we know it is the first in its kind, especially on the application of \gls{svdd} and \gls{svcpd} to real-world \gls{har} data.
The setup follows the structure as illustrated in \Cref{fig:method_overview} and further discussed in \Cref{sec:method_overview}.
It starts with the data collection phase, in which we both use artificial and real-world data.
The data sets are discussed in \Cref{sec:data_gathering}.
After (optional) pre-processing, the \gls{svdd} algorithm is used to create a model representation, which is iteratively updated.
In \Cref{sec:method_model_construction} this construction phase is discussed and shows the windowed approach.
From this constructed model we can extract some features.
In our case, we use the approximated radius $R$ of the hypersphere to detect changes.
\Cref{sec:method_model_features} goes into detail about the rationale of using that feature.
Finally, in \Cref{sec:method_change_detection} we discuss two different, but simple, methods to detect changes in the time series data, based on the extracted radius $R$.

### 2014-01-17 Chapters/Chapter6/results.tex
Move results Chapter 6 to Appendix'

% !TEX root = ../../main.tex
\section{Results}\label{sec:real_world_results}
In this section we will provide the results of our methods \gls{ocs-hats} applied to the real-world data sets and discuss the quality of the solution.
As stated above, we will use the objective quality metrics as used in \Cref{sec:artificial_data_quality_metrics}.
That will be in the form of a tabular overview of the (manually set) best meta-parameters and results, expressed in \gls{far} $\gamma$ and average delay $d_{avg}$ (following \Cref{sec:artificial_data_results}).
Furthermore, we will discuss the results from an empirical point of view, since the problem of finding change points in human activities is ambiguous.
By stating our observations and remarks and providing fragments of visual results, we will discuss the characterizing strengths and weaknesses of our method.

\subsection{Objective metrics}
The employed two-stage algorithm requires the setting of user defined parameters.
In \Cref{tab:results_real_world} for each run the results are listed, together with the used parameters.
We have used a window length of 50 frames, which represents around $1.16$ seconds of data (the frame-rate of the recordings is not fixed).
The other parameter values, \ie the \gls{rbf} $\sigma$, high and low threshold, and proximity parameter $\delta$, are the result of an educated guess and manual optimization.
In general, a high value for the upper threshold and low value for the lower threshold indicates a robust method, since it indicates that change points are significantly different from homogeneous segments.
The proximity $\delta$ should be low, since that indicates there are few false positives.

\begin{table}
  \centering
  \caption[Results real world runs]{Parameter settings and results of the real-world data sets.}
  \begin{tabulary}{\textwidth}{|l|c|c|c|c|c|c|c|}
    \cline{2-8}
    \multicolumn{1}{l|}{} & Run 1 & Run 2 & Run 3 & Run 4 & Run 5 & Run 6 & Run 7 \\
    \hline
    Window length & 50 & 50 & 50 & 50 & 50 & 50 & 50 \\
    \hline
    Sigma of \gls{rbf} & 13 & 13 & 13 & 13 & 13 & 4 & 13 \\
    \hline
    High threshold & 1.2 & 1.5 & 1.1 & 1.3 & 1.2 & 1.7 & 1.3 \\
    \hline
    Low threshold & 0.8 & 0.8 & 0.7 & 0.7 & 0.8 & 0.6 & 0.6 \\
    \hline
    $\delta$ (s) & 0.7 & 0.85 & 0.6 & 0.7 & 1 & 0.8 & 1 \\
    \hline
    \hline
    $\gamma(Y)$ & 0.1 & 0.05 & 0 & 0.2 & 0 & 0.48 & 0 \\
    \hline
    $d_{avg}$ & 0.52 & 0.83 & 0.92 & 1.26 & 1.05 & 0.92 & 0.47 \\
    \hline
    $std(d_{avg})$ & 0.29 & 0.55 & 0.75 & 1.67 & 0.78 & 0.92 & 0.52 \\
    \hline
  \end{tabulary}
  \label{tab:results_real_world}
\end{table}

To give a better impression on the quality of the $\operatorname*{Average\_delay}$, \Cref{fig:boxplot_real_world_runs} shows a box plot for each run.
It shows the spread of the $\operatorname*{Average\_delay}$ in seconds over all the change points.
A lower and more compact box plot is better.

\begin{figure}
\centering
  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/boxplot_results_real_world_runs.eps}
  \caption[Box plot results real-world runs]{Box plot of the results for the real-world runs, indicating the number of data points between the actual and closest detected change points. A lower and more compact box plot is better.}
  \label{fig:boxplot_real_world_runs}
\end{figure}

\subsection{Subjective observations and remarks}
It can be ambiguous where to place the (manually annotated) change points, in the case of real-world data.
The objective quality measure can give a distorted view on the performance.
To overcome this, we will state observations and remarks found after close (visual) inspection of the annotated and discovered change points.
In \Cref{fig:plots_all_runs_results} all the results are plotted.
The black solid lines indicate our manually determined change points.
The dashed purple lines indicate the change points as discovered by our method.
We will start with general observations applicable to all the runs and will continue with individual remarks for each run.
Where possible, a graphical illustration will support our observations.

@@

When looking at the individual runs, we have the following observations for each run number:
\begin{enumerate}
  \item \textbf{Subject 2, straight walk and run} \Cref{fig:data_with_cps_run_1}
    \begin{itemize}
      \item The first transition from running to walking, around $14s$ is harder to discover than the transition for the same activities around $33s$.
      This shows is that in real-world applications there is a diversity between the same transitions and activities.
      \item Around $8s$, $9s$, and $16s$ a few steps (while walking) are regarded as change points.
      During the running segments from $17s$ and $28s$ there is a lower probability of change for each step.
    \end{itemize}
  \item \textbf{Subject 1, straight walk and run} \Cref{fig:data_with_cps_run_2}
    \begin{itemize}
      \item Following the video recordings, we have annotated a change point from running to walking around $37s$.
      Our method discovers a change point almost a second before.
      In retrospect, we can see that the data distribution indeed changes from the discovered change point on.
      This shows us two important principles.
      The first is that the annotations are very subjective.
      The second is that between different activities the transition period is longer than we would think.
      Looking at the data, we can see that the body slows down, even before we visually notice it on the video recordings.
    \end{itemize}

  \item \textbf{Subject 2, walk and run around fountain} \Cref{fig:data_with_cps_run_3}
    \begin{itemize}
      \item Like in the first run, the walking segment from $24s$ results in a change point for each step.
      Further inspection of the data reveals that each step is indeed different from the other.
      Due to the global parameter settings, the sensitivity is too high for this segment to recognize it as one.
      It could also be that our used window-length is too small, to model a good representation of this segment.
      \item During the circular run, from $12s$ till $24$, there are two change points discovered.
      The difference for the rotational vectors need to accumulate to a certain value before they have enough influence to let the rotation be regarded as a change point.
    \end{itemize}
  \item \textbf{Subject 1, walk and run around fountain} \Cref{fig:data_with_cps_run_4}
    \begin{itemize}
      \item As with the other runs, the accelerometer data makes it harder to detect turns.
      It requires a higher sensitivity, which results in a higher \gls{far}.
    \end{itemize}
  \item \textbf{Subject 2, walk and run fountain 2} \Cref{fig:data_with_cps_run_5}
    \begin{itemize}
      \item During the standing segment around $38s$ there are a lot of false positives.
      It seems to be the \emph{false heterogeneity} problem described above.
    \end{itemize}
  \item \textbf{Subject 3, indoor stairs} \Cref{fig:data_with_cps_run_6}
    \begin{itemize}
      \item The walking segment around $22s$, between two segments of walking downstairs, shows little difference in the data.
      To recognize it as a change point a high sensitivity and low proximity time period $\delta$ is required.
      \item During some segments (downstairs from $24s$, upstairs from $42s$ and $54s$) the method discovers more change points than our annotation.
      A closer inspecting of the raw data reveals indeed changes in behavior.
      To exclude these (semi) false positives, a better tuning of parameters is required.
      \item Because of the circular shape of the stairs, the magnetic field sensors constantly differs.
      Although our method is build to exclude slow shifting changes (because we are only interested in sudden changes), with our used window width it still eventually results in change points.
      \item The difference between taking the stairs and walking is smaller than, \eg, walking and running.
      The delay between these segments seems to be larger, as illustrated around $33s$.
    \end{itemize}
  \item \textbf{Subject 2, walk around corner} \Cref{fig:data_with_cps_run_7}
    \begin{itemize}
      \item The $90^{\circ}$ counter-clockwise turn during the walking activity is hard to discover when the accelerometer sensor data is included.
      When only the magnetic field and rotation sensors are used, the turn requires a lower sensitivity.
      With only these two sensors all the other change points in this run are also successfully discovered.
    \end{itemize}
\end{enumerate}

\begin{figure}
  \centering
  \begin{subfigure}{1\textwidth}
    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/run-1-walk-run-roemer/data_plot_acc_with_discovered_cps.eps}
    \caption{Run 1}
    \label{fig:data_with_cps_run_1}
  \end{subfigure}

  \begin{subfigure}{1\textwidth}
    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/run-2-walk-run-jos/data_plot_acc_with_discovered_cps.eps}
    \caption{Run 2}
    \label{fig:data_with_cps_run_2}
  \end{subfigure}

  \begin{subfigure}{1\textwidth}
    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/run-4-run-fountain-roemer/data_plot_acc_with_discovered_cps.eps}
    \caption{Run 3}
    \label{fig:data_with_cps_run_3}
  \end{subfigure}

  \begin{subfigure}{1\textwidth}
    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/run-5-run-fountain-jos/data_plot_acc_with_discovered_cps.eps}
    \caption{Run 4}
    \label{fig:data_with_cps_run_4}
  \end{subfigure}

  \begin{subfigure}{1\textwidth}
    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/run-6-walk-run-roemer/data_plot_acc_with_discovered_cps.eps}
    \caption{Run 5}
    \label{fig:data_with_cps_run_5}
  \end{subfigure}

  \begin{subfigure}{1\textwidth}
    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_acc_with_discovered_cps.eps}
    \caption{Run 6}
    \label{fig:data_with_cps_run_6}
  \end{subfigure}

  \begin{subfigure}{1\textwidth}
    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/run-3-walk-turn-roemer/data_plot_acc_with_discovered_cps.eps}
    \caption{Run 7}
    \label{fig:data_with_cps_run_7}
  \end{subfigure}

  \caption[Results run 1/8]{Annotated plots of all runs. The black vertical lines are manually determined change points, the purple dashes lines are the discovered change points. The label above each segment indicates the current performed activity.}\label{fig:plots_all_runs_results}
\end{figure}

>>>>

% !TEX root = ../../main.tex
\section{Results}\label{sec:real_world_results}
In this section we will provide the results of our methods \gls{ocs-hats} applied to the real-world data sets and discuss the quality of the solution.
As stated above, we will use the objective quality metrics as used in \Cref{sec:artificial_data_quality_metrics}.
That will be in the form of a tabular overview of the (manually set) best meta-parameters and results, expressed in \gls{far} $\gamma$ and average delay $d_{avg}$ (following \Cref{sec:artificial_data_results}).
Furthermore, we have anaylized the results from an emperical point of view, since the problem of finding change points in human activities is ambiguous.
For the sake of brevity we excluded those observations from this chapter and instead xthey can be found in \Cref{AppendixA}.

\subsection{Objective metrics}
The employed two-stage algorithm requires the setting of user defined parameters.
In \Cref{tab:results_final_real_world} for each run the results are listed.
The right-most column summarizes the overall performance by taking the average over all the runs.
We have used a window length of 50 frames, which represents around $1.16$ seconds of data (the frame-rate of the recordings is not fixed).
The other parameter values, \ie the \gls{rbf} $\sigma$, high and low threshold, and proximity parameter $\delta$, are the result of an educated guess and manual optimization.
In general, a high value for the upper threshold and low value for the lower threshold indicates a robust method, since it indicates that change points are significantly different from homogeneous segments.
The proximity $\delta$ should be low, since that indicates there are few false positives.

To give a better impression on the quality of the $d_{avg}$, \Cref{fig:boxplot_final_real_world_runs} shows a box plot for each run.
It shows the spread of the $d_{avg}$ in seconds over all the change points.
A lower and more compact box plot is better.

\begin{table}
  \centering
  \caption[Results real world runs]{Results of the real-world data sets. The used parameters are: window length = $50$, $\sigma = 13$, $th_{high} = 1.2$, $th_{low} = 0.8$}
  \begin{tabulary}{\textwidth}{|l|c|c|c|c|c|c|c||c|}
    \cline{2-9}
    \multicolumn{1}{l|}{} & Run 1 & Run 2 & Run 3 & Run 4 & Run 5 & Run 6 & Run 7 & Average \\
    \hline
    $\gamma(Y)$ & 0.1 & 0.2 & 0.14 & 0.37 & 0.29 & 0.64 & 0.57 & 0.33 \\
    \hline
    $d_{avg}$ & 0.52 & 0.85 & 0.93 & 0.54 & 0.69 & 0.51 & 0.32 & 0.62\\
    \hline
    $std(d_{avg})$ & 0.71 & 0.31 & 0.68 & 0.37 & 0.50 & 0.44 & 0.35 & 0.48 \\
    \hline
  \end{tabulary}
  \label{tab:results_final_real_world}
\end{table}

\begin{figure}
\centering
  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/boxplot_final_runs.eps}
  \caption[Box plot results real-world runs]{Box plot of the results for the real-world runs, indicating the number of data points between the actual and closest detected change points. A lower and more compact box plot is better.}
  \label{fig:boxplot_final_real_world_runs}
\end{figure}

\subsection{Subjective observations and remarks}\label{subsec:subjective_results}
It can be ambiguous where to place the (manually annotated) change points, in the case of real-world data.
The objective quality measure can give a distorted view on the performance.
To overcome this, we will state observations and remarks found after close (visual) inspection of the annotated and discovered change points.
In \Cref{fig:plots_all_runs_results} all the results are plotted.
The black solid lines indicate our manually determined change points.
The dashed purple lines indicate the change points as discovered by our method.
We will start with general observations applicable to all the runs.
A more detailed discussion for each run can be found in \Cref{AppendixA}.
Where possible, a graphical illustration will support our observations.

@@

An more elaborate discussion with observations and remarks is provided in \Cref{AppendixA}.

### 2014-01-17 Chapters/Chapter7.tex
Move results Chapter 6 to Appendix'

In this research we have applied \gls{occ} methods to accelerometer data, recorded during the performance of various human activities, in order to create a temporal segmentation of those recordings.
The goal was to find the change points between activities, assuming that it will aid our understanding of the sensor values.
To do so, we used the \gls{svdd} algorithm by Tax~\cite{tax1999support}, following the method of applying it to a sliding window of data, as done by Camci~\cite{camci2010change}.
Instead of using both the information whether a new data point is an outlier and the radius of the hypersphere to indicate change, we only used the latter model properties.
The increase of the radius of the constructed hypersphere, which encloses the data objects and increases with heterogeneous segments of data, is extracted and used for indication of change.

@@

We have constructed \gls{ocs-hats} as a slightly simplification of \gls{svcpd} by Camci~\cite{camci2010change} application to real-world inertial sensor data.
We have tested our proposed method with the same artificial data and did not find a deterioration in the results.

@@

Finally, due to encountered problems with commonly used data sets \cite{kwapisz2011activity,anguita2012human}, we have recorded our real-world data sets by ourselves.
We have recorded all the activities with a smartphone worn in the right front pocket of the subjects.
Furthermore, we have recorded the performing subjects from the third person point of view with a video camera.
This allows for better interpretation of the results.
All the recorded data is made publicity available under the \gls{almende-data} data set \cite{vlasveld2014acras}.

\TODO{Vraag aan Anne: toevoegen samenvatting van resultaten uit \Cref{Chapter6}: de subjectieve waarnemingen?}

\section{Limitations}\label{sec:limitations}
Our proposed method was constructed under the assumption that an explicit temporal segmentation of inertial sensor data can aid the process of activity classification.
We have expressed the problem of temporal segmentation as finding change points in the time series data.
Since this is the first application of \gls{oc-svm} based algorithms to real-world accelerometer data, we have limited ourselves to creating a first method.
The (results of the) proposed method is subject to parameters, which we have set manually based on empirical results and experience.

We have applied our method to real-world data, recorded in 7 runs.
For a better and more robust analysis of the algorithm it can be applied to more real-world data.
Although we have recorded and publicized our data, application to other commonly used data sets would enable a fair comparison.
Furthermore, since the manual annotation of change points is an ambiguous task, the results remain subjective.

@@

\section{Final words}\label{sec:final_remarks}
In this research we have constructed a method to segment inertial sensor time series data, from performed human activities.
To this end we have applied \gls{occ} and \gls{oc-svm} based method to our own recorded data sets.
We followed the approach by Camci~\cite{camci2010change} and used the theory and implementation of \gls{svdd} by Tax~\cite{tax2001one,Ddtools2013}.
Although the performance of the method heavily depends on user set parameters, the results are promising.
We believe that the temporal segmentation of time series data can aid the process of activity classification, which is a logical follow-up on this research.
\TODO{Vraag aan Anne: Hier nog meer samenvatten? Wat is een goede uitsmijter?}

>>>>

This research has applied \gls{occ} methods to inertial sensor data, recorded during the performance of various human activities, in order to create a temporal segmentation of those recordings.
The goal was to find the change points between activities, assuming that it will aid our understanding of the sensor values.
To do so, we used the \gls{svdd} algorithm by Tax~\cite{tax1999support}, following the method of applying it to a sliding window of data, as done by Camci~\cite{camci2010change}.
From these methods we have created our own method \gls{ocs-hats} and made it puclibly available \cite{vlasveld2014hats}, which is a slightly modified implemention of \gls{svcpd} by Camci~\cite{camci2010change} applied to \gls{har} data.
% Instead of using both the information whether a new data point is an outlier and the radius of the hypersphere to indicate change, we only used the latter model properties.
An increase of the radius of the constructed hypersphere, which encloses the data objects and increases with heterogeneous segments of data, is extracted and used for indication of change.

@@

We have constructed \gls{ocs-hats} as a slightly simplification of \gls{svcpd} by Camci~\cite{camci2010change} application to real-world inertial sensor data.
The \textsc{Matlab} implementation is free to use for the public \cite{vlasveld2014hats}.
We have tested our proposed method with the same artificial data and did not find a deterioration in the results.
Because of the noisy nature of the recorded activities and the ability to segment the time series, we have shown that \gls{ocs-hats} is a robust method for real-world applications.

@@

The proposed \gls{ocs-hats} method processes the raw inertial sensor data, originating from sensors as accelerometers and magnetic field sensors.
This in contrast with many other approaches that extract features in a pre-processing stage.
The method processes the data in an online manner, which makes it suitable for applications that need direct segmentation of time series data.

Finally, due to encountered problems with commonly used data sets \cite{kwapisz2011activity,anguita2012human}, we have recorded our real-world data sets by ourselves.
We have recorded all the activities with a smartphone worn in the right front pocket of the subjects.
Furthermore, we have recorded the performing subjects from the third person point of view with a video camera.
This allows for better interpretation of the results, since it provides a ground truth which can be consulted in case of ambiguity.
All the recorded data is made publicly available as the \gls{almende-data} data set \cite{vlasveld2014acras}.
Using our results as a benchmark, it is possible for others to use our data set and improve on it.

\section{Limitations}\label{sec:limitations}
Our proposed method \gls{ocs-hats} was constructed under the assumption that an explicit temporal segmentation of inertial sensor data can aid the process of activity classification.
We have expressed the problem of temporal segmentation as finding change points in the time series data.
Since this is the first application of \gls{oc-svm} based algorithms to real-world accelerometer data, we have limited ourselves to creating an initial method, without optimization.
The (results of the) proposed method is subject to parameters, which we have set manually based on empirical results and experience.

We have applied our method to real-world data, recorded in 7 runs.
For a better and more robust analysis of the algorithm it can be applied to other real-world data sets.
Although we have recorded and publicized our data, it would be nice if others also record continuous data sets with ground truth labeling, for fair comparison.
However, due to the application domain and ambiguity, the results remain subjective.

For the implementation of \gls{ocs-hats} we have created and used a \textsc{Matlab} application.
Because of that, we can not make reliable claims on the speed and memory usage of the algorithm.
With the online nature and applications in mind and different targeted hardware platforms, this is something that could be of importance.

@@

\section{Final words}\label{sec:final_remarks}
In this research we have constructed a method to segment inertial sensor time series data, from performed human activities.
To this end we have applied \gls{occ} and \gls{oc-svm} based method to our own recorded data sets.
We followed the approach by Camci~\cite{camci2010change} and used the theory and implementation of \gls{svdd} by Tax~\cite{tax2001one,Ddtools2013}.
Although the performance of the method heavily depends on user set parameters, the results are promising.
We believe that the temporal segmentation of time series data can aid the process of activity classification, which is a logical follow-up on this research.

We have created a continuous recorded data set \gls{almende-data}~\cite{vlasveld2014acras} from inertial sensors found in typical modern smartphones.
The transition periods between the activities are part of the time series data.
This makes the data set useful for research on segmentation.
Besides the manually annotated change points can the ground truth of the performed activity be determined by using the provided video recordings of the activities.

With our method for temporal segmentation of human activities, \gls{ocs-hats}~\cite{vlasveld2014hats}, we believe to have contributed to the field of machine learning in the context of \acrlong{har}.
